### 2025-02-06 paper/body_ecml.tex
uh

\subsection{Encoding Domain Knowledge}\label{encoding-domain-knowledge}

>>>>

\subsection{Encoding Actionability
Constraints}\label{encoding-actionability-constraints}

Many existing counterfactual explainers support domain and mutability
constraints out-of-the-box. In fact, both types of constraints can be
implemented for any counterfactual explainer that relies on gradient
descent in the feature space for optimization
\cite{altmeyer2023explaining}. In this context, domain constraints can
be imposed by simply projecting counterfactuals back to the specified
domain, if the previous gradient step moved resulted in updated feature
values out of the domain. Mutability constraints can similarly be
enforced by simply clipping gradients to ensure that features are only
mutated in the allowed direction, if at all.

Since both types of constraints directly affect \(\mathbf{x}^\prime\),
they also influence model outcomes through Equation~\ref{eq-div}.

Mutability constraints

### 2025-02-06 paper/preprint/paper.tex
uh

\subsection{Encoding Domain Knowledge}\label{encoding-domain-knowledge}

@@

>>>>

\subsection{Encoding Actionability
Constraints}\label{encoding-actionability-constraints}

Many existing counterfactual explainers support domain and mutability
constraints out-of-the-box. In fact, both types of constraints can be
implemented for any counterfactual explainer that relies on gradient
descent in the feature space for optimization
(\citeproc{ref-altmeyer2023explaining}{Altmeyer, Deursen, et al. 2023}).
In this context, domain constraints can be imposed by simply projecting
counterfactuals back to the specified domain, if the previous gradient
step moved resulted in updated feature values out of the domain.
Mutability constraints can similarly be enforced by simply clipping
gradients to ensure that features are only mutated in the allowed
direction, if at all.

Since both types of constraints directly affect \(\mathbf{x}^\prime\),
they also influence model outcomes through Equation~\ref{eq-div}.

Mutability constraints

@@

\bibitem[\citeproctext]{ref-altmeyer2023explaining}
Altmeyer, Patrick, Arie van Deursen, et al. 2023. {``Explaining
Black-Box Models Through Counterfactuals.''} In \emph{Proceedings of the
JuliaCon Conferences}, 1:130. 1.

### 2025-02-06 paper/sections/method.qmd
vamos
uh

## Encoding Domain Knowledge

>>>>

## Encoding Actionability Constraints

Many existing counterfactual explainers support domain and mutability constraints out-of-the-box. In fact, both types of constraints can be implemented for any counterfactual explainer that relies on gradient descent in the feature space for optimization [@altmeyer2023explaining]. In this context, domain constraints can be imposed by simply projecting counterfactuals back to the specified domain, if the previous gradient step moved resulted in updated feature values out of the domain. Mutability constraints can similarly be enforced by simply clipping gradients to ensure that features are only mutated in the allowed direction, if at all.

Since both types of constraints directly affect $\mathbf{x}^\prime$, they also influence model outcomes through @eq-div.

Mutability constraints
